Dominance Regions for Imaginary $\bfg$-vectors

### 2020-06-19 dominance_regions.tex
Lemma 3.7 is proved, I was expecting to need a change for i odd but this does not seem to be the case, some cleanup is needed but it is late

  Observe that $\phi_{t_1}$ is the piecewise-linear map given by
  \[\phi_{t_1}(\lambda)=\begin{cases} (-\lambda_1,c\lambda_1+\lambda_2) & \text{if $\lambda_1\ge0$;}\\ (-\lambda_1,\lambda_2) & \text{if $\lambda_1<0$;} \end{cases}\]
  and $\phi_{t_2}=\phi_{t_{-2}}^{-1}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_{t_2}(\lambda)=\begin{cases} ((bc-1)\lambda_1+b\lambda_2, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$;}\\ (-\lambda_1, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2<0$;}\\ (-\lambda_1+b\lambda_2, -\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2\ge 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2<0$;}\end{cases}
  \end{equation}
  while $\phi_{t_2}^{-1}=\phi_{t_{-2}}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:backward two step mutation}
    \phi_{t_2}^{-1}(\lambda)=\begin{cases} (-\lambda_1-b\lambda_2, c\lambda_1+(bc-1)\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2\le 0$;}\\ (-\lambda_1-b\lambda_2, -\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2>0$;}\\ (-\lambda_1, c\lambda_1-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1\le 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1>0$;}\end{cases}
  \end{equation}

@@

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  Also, write $\cI' \subset \RR^2$ for the \emph{dual imaginary cone} spanned by the vectors $(-bc\pm\sqrt{bc(bc-4)},2c)$.
  \begin{lemma}
    The vectors $(2b, -bc+\sqrt{bc(bc-4)})$ (resp. $(2b, -bc-\sqrt{bc(bc-4)})$) and $(2c, bc+\sqrt{bc(bc-4)})$ (resp. $(2c, bc-\sqrt{bc(bc-4)})$) are perpendicular.
  \end{lemma}
  \begin{proof}
    This is immediate from the equality $(-bc+\sqrt{bc(bc-4)})(bc+\sqrt{bc(bc-4)})=-4bc$.
  \end{proof}
  \begin{lemma}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have $\phi_{t_{2j}}(\lambda)\in\cI$ and $\phi_{t_{2j+1}}(\lambda)\in\cI'$.
  \end{lemma}
  \begin{proof}
    Note that $\phi_{t_{2j}}=\phi_{t_2}^j$ and $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$ so the result follows from the cases $j=1$ and $j=0$ respectively.
    The claim is immediate for $\phi_{t_2}$ since the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$ are eigenvectors.
    Now, we have $\phi_{t_1}(\lambda)=(-\lambda_1,c\lambda_1+\lambda_2)$ and so $\phi_{t_1}\big( (2b,-bc\pm\sqrt{bc(bc-4)}) \big)=(-2b,bc\pm\sqrt{bc(bc-4)})$, which is the $\frac{bc\pm\sqrt{bc(bc-4)}}{2c}$-multiple of the boundary vector for $\cI'$ and the result follows.
  \end{proof}

@@

  \begin{lemma}
    For $i\ge2$ and $\lambda\in\cI$, the region $\phi_{t_i}^{-1}\cC_{t_i}(\phi_{t_i}\lambda)\subset\RR^2$ is the polytope determined by the following inequalities:
    \begin{align}
      \label{ineq:1} u_{i,-}x+u_{i-1,+}y &\le u_{i,-}\lambda_1+u_{i-1,+}\lambda_2;\\
      \label{ineq:2} u_{i+1,-}x+u_{i,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2;\\
      \label{ineq:3} -u_{i-1,-}x+u_{i,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2;\\
      \label{ineq:4} -u_{i-1,-}x-u_{i-2,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2.
    \end{align}

@@

    Assuming the inequalities (a-d) hold for $k$, we apply Lemma~\ref{le:transformed inequalities} with $\phi_{t_2}^{-1}$.
    Both of the boundary rays corresponding to the inequalities (a) and (d) lies entirely in the region where $\phi_{t_2}^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & -b\\ c & bc-1 \end{array}\right]$.
    Thus applying Lemma~\ref{le:transformed inequalities} to (a) gives the inequality 
    \[(u_{3,+}u_{2k,-}-u_{2,-}u_{2k-1,+})x+(u_{2,+}u_{2k,-}-u_{1,-}u_{2k-1,+})y\le u_{i,-}\lambda_1+u_{i-1,+}\lambda_2,\]
    which is the inequality (a) for $k+1$ by Lemma~\ref{eq:long Chebyshev recursion}, while applying this to (d) gives the inequality 
    \[(-u_{3,+}u_{2k-1,-}+u_{2,-}u_{2k-2,+})x+(-u_{2,+}u_{2k-1,-}+u_{1,-}u_{2k-2,+})y\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2,\]
    which is the inequality (d) for $k+1$ again by Lemma~\ref{eq:long Chebyshev recursion}.
    Similarly, the boundary segment corresponding to (b) lies entirely in the region where $\phi_{t_2}^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ c & -1 \end{array}\right]$.
    Thus applying Lemma~\ref{le:transformed inequalities} to (c) gives the inequality 
    \[(u_{1,+}u_{2k-1,-}-u_{2,-}u_{2k,+})x+(-u_{0,+}u_{2k-1,-}-u_{1,-}u_{2k,+})y\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2,\]
    which is the inequality (d) for $k+1$ by Lemma~\ref{eq:long Chebyshev recursion}, in particular we see that the segment determined by (c) and the ray determined by (d) align in the image.

    For $i=2j+1$, $j>0$, and $\lambda\in\cI$, we get $\cC_{t_i}(\phi_{t_i}\lambda)\subset\RR^2$ is given by the inequalities 
    \[ -x\le u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2 \qquad\text{and}\qquad y\le u_{2j+2,-}\lambda_1+u_{2j+1,-}\lambda_2. \]
  \end{proof}

>>>>

  Observe that $\phi_{t_1}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward mutation 1}
    \phi_{t_1}(\lambda)=\begin{cases} (-\lambda_1,c\lambda_1+\lambda_2) & \text{if $\lambda_1\ge0$;}\\ (-\lambda_1,\lambda_2) & \text{if $\lambda_1<0$;} \end{cases}
  \end{equation}
  and $\phi_{t_2}=\phi_{t_{-2}}^{-1}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_{t_2}(\lambda)=\begin{cases} ((bc-1)\lambda_1+b\lambda_2, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$;}\\ (-\lambda_1, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2<0$;}\\ (-\lambda_1+b\lambda_2, -\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2\ge 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2<0$;}\end{cases}
  \end{equation}
  while $\phi_{t_2}^{-1}=\phi_{t_{-2}}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:backward two step mutation}
    \phi_{t_2}^{-1}(\lambda)=\begin{cases} (-\lambda_1-b\lambda_2, c\lambda_1+(bc-1)\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2\le 0$;}\\ (-\lambda_1-b\lambda_2, -\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2>0$;}\\ (-\lambda_1, c\lambda_1-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1\le 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1>0$;}\end{cases}
  \end{equation}

@@

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  Also, write $\cI' \subset \RR^2$ for the \emph{dual imaginary cone} spanned by the vectors $(-bc\pm\sqrt{bc(bc-4)},2c)$.
  \begin{lemma}
    The vectors $(2b, -bc+\sqrt{bc(bc-4)})$ (resp. $(2b, -bc-\sqrt{bc(bc-4)})$) and $(2c, bc+\sqrt{bc(bc-4)})$ (resp. $(2c, bc-\sqrt{bc(bc-4)})$) are perpendicular.
  \end{lemma}
  \begin{proof}
    This is immediate from the equality $(-bc+\sqrt{bc(bc-4)})(bc+\sqrt{bc(bc-4)})=-4bc$.
  \end{proof}
  \begin{lemma}
    \label{le:imaginary stability}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have $\phi_{t_{2j}}(\lambda)\in\cI$ and $\phi_{t_{2j+1}}(\lambda)\in\cI'$.
  \end{lemma}
  \begin{proof}
    Note that $\phi_{t_{2j}}=\phi_{t_2}^j$ and $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$ so the result follows from the cases $j=1$ and $j=0$ respectively.
    The claim is immediate for $\phi_{t_2}$ since the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$ are eigenvectors.
    Now, we have $\phi_{t_1}(\lambda)=(-\lambda_1,c\lambda_1+\lambda_2)$ and so $\phi_{t_1}\big( (2b,-bc\pm\sqrt{bc(bc-4)}) \big)=(-2b,bc\pm\sqrt{bc(bc-4)})$, which is the $\frac{bc\pm\sqrt{bc(bc-4)}}{2c}$-multiple of the boundary vector for $\cI'$ and the result follows.
  \end{proof}

@@

  \begin{lemma}
    For $i\ge1$ and $\lambda\in\cI$, the region $\phi_{t_i}^{-1}\cC_{t_i}(\phi_{t_i}\lambda)\subset\RR^2$ is the polytope determined by the following inequalities:
    \begin{align}
      \label{ineq:1} u_{i,-}x+u_{i-1,+}y &\le u_{i,-}\lambda_1+u_{i-1,+}\lambda_2;\\
      \label{ineq:2} u_{i+1,-}x+u_{i,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2;\\
      \label{ineq:3} -u_{i-1,-}x+u_{i,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2;\\
      \label{ineq:4} -u_{i-1,-}x-u_{i-2,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2.
    \end{align}

@@

    Assuming the inequalities (a)-(d) hold for $k$, we apply Lemma~\ref{le:transformed inequalities} with $\phi_{t_2}^{-1}$.
    Both of the boundary rays corresponding to the inequalities (a) and (d) lies entirely in the region where $\phi_{t_2}^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & -b\\ c & bc-1 \end{array}\right]$, also the boundary segment corresponding to (b) intersects this region.
    Thus applying Lemma~\ref{le:transformed inequalities} to (a) gives the inequality 
    \[(u_{3,+}u_{2k,-}-u_{2,-}u_{2k-1,+})x+(u_{2,+}u_{2k,-}-u_{1,-}u_{2k-1,+})y\le u_{i,-}\lambda_1+u_{i-1,+}\lambda_2,\]
    which is the inequality (a) for $k+1$ by Lemma~\ref{eq:long Chebyshev recursion}; while applying this to (d) gives the inequality 
    \[(-u_{3,+}u_{2k-1,-}+u_{2,-}u_{2k-2,+})x+(-u_{2,+}u_{2k-1,-}+u_{1,-}u_{2k-2,+})y\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2,\]
    which is the inequality (d) for $k+1$ again by Lemma~\ref{eq:long Chebyshev recursion}; finally applying this to (b) gives the inequality 
    \[(u_{3,+}u_{2k+1,-}-u_{2,-}u_{2k,+})x+(u_{2,+}u_{2k+1,-}-u_{1,-}u_{2k,+})y\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2,\]
    which is the inequality (b) for $k+1$
    .
    Similarly, the boundary segment corresponding to (c) lies entirely in the region where $\phi_{t_2}^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ c & -1 \end{array}\right]$.
    Thus applying Lemma~\ref{le:transformed inequalities} to (c) gives the inequality 
    \[(u_{1,+}u_{2k-1,-}-u_{2,-}u_{2k,+})x+(-u_{0,+}u_{2k-1,-}-u_{1,-}u_{2k,+})y\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2,\]
    which is the inequality (d) for $k+1$ by Lemma~\ref{eq:long Chebyshev recursion}, in particular we see that the segment determined by (c) and the ray determined by (d) align in the image.
    Lastly, the boundary segment corresponding to (b) also intersects the regions where $\phi_{t_2}^{-1}$ acts according to the matrices $\left[ \begin{array}{cc} -1 & -b\\ 0 & -1 \end{array}\right]$ and $\left[ \begin{array}{cc} -1 & 0\\ 0 & -1 \end{array}\right]$ respectively.
    Applying Lemma~\ref{le:transformed inequalities} to (b) with the first matrix gives the inequality 
    \[(-u_{1,+}u_{2k+1,-}+u_{0,-}u_{2k,+})x+(u_{2,+}u_{2k+1,-}-u_{1,-}u_{2k,+})y\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2,\]
    which is the inequality (c) for $k+1$ by Lemma~\ref{eq:long Chebyshev recursion}, while applying Lemma~\ref{le:transformed inequalities} to (b) with the second matrix gives the inequality 
    \[-u_{2k+1,-}x-u_{2k,+}y\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2,\]
    which again reproduces the inequality (d) and aligns with the previous segment and ray in the image.

    For $i=2j+1$, $j>0$, and $\lambda\in\cI$, we get $\cC_{t_{2j+1}}(\phi_{t_{2j+1}}\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger')\ -x\le u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2 \qquad\text{and}\qquad (\ddagger')\ y\le u_{2j+2,-}\lambda_1+u_{2j+1,-}\lambda_2.\]
    Using that $\phi_{t_{2j+1}}^{-1}=\phi_{t_2}^{-j}\phi_{t_1}^{-1}$, we compute the image inductively as above.
    From \eqref{eq:forward mutation 1} and Lemma~\ref{le:imaginary stability}, we see that the boundary ray for $\cC_{t_{2j+1}}(\phi_{t_{2j+1}}\lambda)$ corresponding to ($\dagger'$) lies entirely in the region in which $\phi_{t_1}^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ c & 1 \end{array}\right]$ of determinant $-1$.
    Thus applying Lemma~\ref{le:transformed inequalities}, the inequality ($\dagger'$) is transformed by $\phi_{t_1}^{-1}$ into the inequality $x\le u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2$.
    The boundary ray corresponding to ($\ddagger'$) intersects both domains of linearity for $\phi_{t_1}^{-1}$ and thus produces the inequalities
    \[ cx+y\le u_{2j+2,+}\lambda_1+u_{2j+1,+}\lambda_2 \qquad\text{and}\qquad  y\le u_{2j+2,+}\lambda_1+u_{2j+1,+}\lambda_2.\]

    \subsubsection*{Claim:} For $k\ge 0$, $\phi_{t_{2k+1}}^{-1}\cC_{t_{2j+1}}(\phi_{t_{2j+1}}\lambda)$ is the region determined by the inequalities 
    \begin{align*}
      \tag{a$'$} u_{2k+1,-}x+u_{2k,+}y &\le u_{i,-}\lambda_1+u_{i-1,+}\lambda_2;\\
      \tag{b$'$} u_{2k+2,-}x+u_{2k+1,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2;\\
      \tag{c$'$} -u_{2k,-}x+u_{2k+1,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2;\\
      \tag{d$'$} -u_{2k,-}x-u_{2k-1,+}y &\le u_{i+1,-}\lambda_1+u_{i,+}\lambda_2.
    \end{align*}
    By essentially the same calculations as above, these inequalities reproduce under the action of $\phi_{t_2}^{-1}$ and this completes the proof.
  \end{proof}
